---
title       : Analyzing College Earnings Data 
subtitle    : Data analysis based on recently released White House report 
author      : Ramesh Subramanian
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, quiz]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

### Context for Data Analysis
The White House recently released a report on the state of College Education in the U.S.
Data collected from 7804 colleges offering a 4 year degree was analyzed along the following four dimensions, six, eight, and ten years after completion of degree.    
1. Academics  2. cost of education  3. Earnings   4. Debt  

#### Sources of Data 
1.  https://collegescorecard.ed.gov
2.   www.collegeboard.edu/collegeScoreCard (7804 rows, 109 variables)

#### This report    
1.  Subset analyzing only 6 year data along dimensions of earnings and debt  
2.  Imputed missing values with median value of column (may introduce some bias)  
3.  Data file for analysis (and on github): collegeEarningsData.csv

--- .class #id 

## Data extraction and imputation

#### Data Extraction and cleansing  
1. Raw data obtained from the web site has 7804 rows and 109 columns.  
2. There are many values that are either NULL or marked as "Privacy Suppressed" indicating that the college did not want to disclose that piece of data.
3.  Only a subset of the data was extracted (7804 rows, 12 columns) because that is the basis for our analysis.

#### Data imputation 
1.  All the values marked as NULL or "Privacy Suppressed" were replaced with NA
2.  All the NA values were replaced with the median value of that column. This may introduce some bias in the analysis.

#### Data Set used for analysis  
1.  earnings.csv (7804 rows, 12 columns)

--- .class #id

## Set up ui.R: Only Pseudocode displayed due to space limitations

#### Set up Help text within sidebarPanel
```{r message=FALSE, eval=FALSE}
   helpText("...Text explaining how application should be used...") 
```

#### Set up selectInput box within sidebarPanel with input choices and default choice
```{r message=FALSE, eval=FALSE}
selectInput("var", 
            label = "Choose data to display",
            choices = list("...list of 5 choices goes here..."),
                  selected = 1)),
```

#### Set up mainPanel within sideBarLayout
```{r message=FALSE, eval=FALSE}
   
   mainPanel(
      h3(textOutput("text1"),align = "center"),
      tableOutput("school_table")
    )
```

--- .class #id

##  Set up server.R: Only Pseudocode displayed due to space limitations

####  Set up data tables for each input choice outside of ShinyServer code    
1.  Earnings (top and bottom 25 colleges)
2.  Percent earning more than $25K ((top and bottom 25 colleges)
3.  Median debt for Low, Medium and High Income families

#### Shiny server code that will 'reactively' display data based on input selection
```{r message=FALSE, eval=FALSE}
    output$school_table <- renderTable({ 
      var <- switch(input$var, 
                     "...choice of input here..." = data_table_to_display)
       })
```

--- .class #id

## Quiz and Conclusions

--- &radio

How many colleges were used in the analysis?

1.  109
2.  _7804_
3.  12
4.  A lot

*** .hint
  This is displayed in the application as well as in this presentation

*** .explanation
  Data from 7804 colleges across the U.S. was used in the analysis.

#### Conclusions  
1.  All of the 25 top colleges had earnings higher than $75K per year
2.  Range of annual earnings for bottom 25 colleges was $7K-$11K
3.  There are many who are making less than $25K per year,  6 years after graduation
4.  Average debt levels seem to be about the same regardles sof income levels implying financial assistance programs pay for most of college tution costs.

#### Fatal flaw: Qualitative aspects of education ignored. One of the pitfalls of drawing conclusions based on data alone.

--- .class #id